### 2020-11-08 Notebooks/intr_results.rmd
Update intr_results.rmd

#install.packages("tidyverse")
#install.packages("COVID19")
library(COVID19)
library(tidyverse)

owid <- read_csv("https://raw.githubusercontent.com/owid/covid-19-data/master/public/data/owid-covid-data.csv")
owid_jan <- owid %>% filter(date=="2020-01-31")
owid_feb <- owid %>% filter(date=="2020-02-29")
owid_mar <- owid %>% filter(date=="2020-03-31")
owid_ap <- owid %>% filter(date=="2020-04-30")
owid_may <- owid %>% filter(date=="2020-05-31")
owid_jun <- owid %>% filter(date=="2020-06-30")
owid_jul <- owid %>% filter(date=="2020-07-31")
owid_ag <- owid %>% filter(date=="2020-08-31")
owid_sep <- owid %>% filter(date=="2020-09-30")
owid_oct <- owid %>% filter(date=="2020-10-31")
owid_list=list(jan20=owid_jan,feb20=owid_feb,march20=owid_mar,april20=owid_ap,may20=owid_may,june20=owid_jun,july=owid_jul,
               ag=owid_ag,sep=owid_sep,oct=owid_oct)

#saving datasets
jan <-covid19(start ="2020-01-31" ,end ="2020-01-31",raw = F ) 
feb <-covid19(start ="2020-02-29" ,end ="2020-02-29",raw = F ) 
march <- covid19(start ="2020-03-31" ,end ="2020-03-31",raw = F ) 
april <-covid19(start ="2020-04-30" ,end ="2020-04-30",raw = F )  
may <- covid19(start ="2020-05-31" ,end ="2020-05-31" ,raw = F) 
june <- covid19(start ="2020-06-30" ,end ="2020-06-30",raw = F ) 
july <- covid19(start ="2020-07-31" ,end ="2020-07-31",raw = F ) 
ag <- covid19(start ="2020-08-31" ,end ="2020-08-31",raw = F ) 
sep <- covid19(start ="2020-09-30" ,end ="2020-09-30",raw = F ) 
oct <- covid19(start ="2020-10-31" ,end ="2020-10-31",raw = F )

#list of datasets
datasets <- list(jan20=jan,feb20=feb,march20=march,april20=april,may20=may,june20=june,july=july,ag20=ag,sep20=sep,oc20=oct)

#renaming a few cells in the datasets

for(i in 1:length(datasets)){

  datasets[[i]]$school_closing <- gsub(0, 0, datasets[[i]]$school_closing)
  datasets[[i]]$school_closing <- gsub(1, 0, datasets[[i]]$school_closing)
  datasets[[i]]$school_closing <- gsub(2, 0, datasets[[i]]$school_closing)
  datasets[[i]]$school_closing <- gsub(3, 1, datasets[[i]]$school_closing)

  datasets[[i]]$workplace_closing <- gsub(0,0, datasets[[i]]$workplace_closing)
  datasets[[i]]$workplace_closing <- gsub(1, 0, datasets[[i]]$workplace_closing)
  datasets[[i]]$workplace_closing <- gsub(2, 0, datasets[[i]]$workplace_closing)
  datasets[[i]]$workplace_closing <- gsub(3, 1, datasets[[i]]$workplace_closing)

  datasets[[i]]$cancel_events<- gsub(0, 0, datasets[[i]]$cancel_events)
  datasets[[i]]$cancel_events <- gsub(1, 0, datasets[[i]]$cancel_events)
  datasets[[i]]$cancel_events <- gsub(2, 1, datasets[[i]]$cancel_events)

  datasets[[i]]$transport_closing<- gsub(0,0, datasets[[i]]$transport_closing)
  datasets[[i]]$transport_closing<- gsub(1, 0, datasets[[i]]$transport_closing)
  datasets[[i]]$transport_closing<- gsub(2, 1, datasets[[i]]$transport_closing)

  datasets[[i]]$stay_home_restrictions<- gsub(0, 0, datasets[[i]]$stay_home_restrictions)
  datasets[[i]]$stay_home_restrictions<- gsub(1, 0, datasets[[i]]$stay_home_restrictions)
  datasets[[i]]$stay_home_restrictions<- gsub(2, 0, datasets[[i]]$stay_home_restrictions)
  datasets[[i]]$stay_home_restrictions<- gsub(3, 1, datasets[[i]]$stay_home_restrictions)

  datasets[[i]]$internal_movement_restrictions<- gsub(0, 0, datasets[[i]]$internal_movement_restrictions)
  datasets[[i]]$internal_movement_restrictions<- gsub(1, 0, datasets[[i]]$internal_movement_restrictions)
  datasets[[i]]$internal_movement_restrictions<- gsub(2, 1, datasets[[i]]$internal_movement_restrictions)

  datasets[[i]]$international_movement_restrictions<- gsub(0, 0, datasets[[i]]$international_movement_restrictions)
  datasets[[i]]$international_movement_restrictions<- gsub(1, 0, datasets[[i]]$international_movement_restrictions)
  datasets[[i]]$international_movement_restrictions<- gsub(2, 0, datasets[[i]]$international_movement_restrictions)
  datasets[[i]]$international_movement_restrictions<- gsub(3, 0, datasets[[i]]$international_movement_restrictions)
  datasets[[i]]$international_movement_restrictions<- gsub(4, 1, datasets[[i]]$international_movement_restrictions)

  names(datasets[[i]])[names(datasets[[i]]) == "iso_alpha_3"] <- "iso_code"

}

for(i in 1:length(datasets)){
  datasets[[i]]<-merge(owid_list[[i]],datasets[[i]],by="iso_code")
}

#School  closures mar world, europe, sa, na
continents=list(w=c(sc=list(),work=list()),eu=c(sc=list(),work=list()),sa=c(sc=list(),work=list()),na=c(sc=list(),work=list()))
for(i in 1:length(datasets)){

continents$w$sc[[i]]=data.frame(table(datasets[[i]]$school_closing))

continents$eu$sc[[i]]=as.data.frame(table(datasets[[i]]$school_closing[datasets[[i]]$continent=="Europe"]))

continents$sa$sc[[i]]=data.frame(table(datasets[[i]]$school_closing[datasets[[i]]$continent=="South America"]))

continents$na$sc[[i]]=data.frame(table(datasets[[i]]$school_closing[datasets[[i]]$continent=="North America"]))

continents$w$work[[i]]=data.frame(table(datasets[[i]]$workplace_closing))

continents$eu$work[[i]]=data.frame(table(datasets[[i]]$workplace_closing[datasets[[i]]$continent=="Europe"]))

continents$sa$work[[i]]=data.frame(table(datasets[[i]]$workplace_closing[datasets[[i]]$continent=="South America"]))

continents$na$work[[i]]=data.frame(table(datasets[[i]]$workplace_closing[datasets[[i]]$continent=="North America"]))

continents$w$events[[i]]=data.frame(table(datasets[[i]]$cancel_events))

continents$eu$events[[i]]=data.frame(table(datasets[[i]]$cancel_events[datasets[[i]]$continent=="Europe"]))

continents$sa$events[[i]]=data.frame(table(datasets[[i]]$cancel_events[datasets[[i]]$continent=="South America"]))

continents$na$events[[i]]=data.frame(table(datasets[[i]]$cancel_events[datasets[[i]]$continent=="North America"]))

#########

continents$w$transp[[i]]=data.frame(table(datasets[[i]]$transport_closing))

continents$eu$transp[[i]]=data.frame(table(datasets[[i]]$transport_closing[datasets[[i]]$continent=="Europe"]))

continents$sa$transp[[i]]=data.frame(table(datasets[[i]]$transport_closing[datasets[[i]]$continent=="South America"]))

continents$na$transp[[i]]=data.frame(table(datasets[[i]]$transport_closing[datasets[[i]]$continent=="North America"]))

#########

continents$w$home[[i]]=data.frame(table(datasets[[i]]$stay_home_restrictions))

continents$eu$home[[i]]=data.frame(table(datasets[[i]]$stay_home_restrictions[datasets[[i]]$continent=="Europe"]))

continents$sa$home[[i]]=data.frame(table(datasets[[i]]$stay_home_restrictions[datasets[[i]]$continent=="South America"]))

continents$na$home[[i]]=data.frame(table(datasets[[i]]$stay_home_restrictions[datasets[[i]]$continent=="North America"]))

#########

continents$w$internal[[i]]=data.frame(table(datasets[[i]]$internal_movement_restrictions))

continents$eu$internal[[i]]=data.frame(table(datasets[[i]]$internal_movement_restrictions[datasets[[i]]$continent=="Europe"]))

continents$sa$internal[[i]]=data.frame(table(datasets[[i]]$internal_movement_restrictions[datasets[[i]]$continent=="South America"]))

continents$na$internal[[i]]=data.frame(table(datasets[[i]]$internal_movement_restrictions[datasets[[i]]$continent=="North America"]))

#########

continents$w$international[[i]]=data.frame(table(datasets[[i]]$international_movement_restrictions))

continents$eu$international[[i]]=data.frame(table(datasets[[i]]$international_movement_restrictions[datasets[[i]]$continent=="Europe"]))

continents$sa$international[[i]]=data.frame(table(datasets[[i]]$international_movement_restrictions[datasets[[i]]$continent=="South America"]))

continents$na$international[[i]]=data.frame(table(datasets[[i]]$international_movement_restrictions[datasets[[i]]$continent=="North America"]))

}

continents$na$international[[10]]