### 2021-01-11 Notebooks/paper_covid.rmd
Update paper_covid.rmd

---
title: "Covid-19 A Spatio-Temporal Investigation."
author: "Lucas Rabelo"
output:
  pdf_document:
    toc: no
    toc_depth: '6'
    number_sections: no
    keep_tex: yes
    citation_package: natbib
    pandoc_args: --listings
    includes: null
  word_document:
    toc: no
    toc_depth: '6'
geometry: margin=1in
fontsize: 12pt
linestretch: 1.5
documentclass: report
---

## Introduction {.tabset .tabset-fade .tabset-pills}

 At the end of 2019, the situation of infections by Covid-19 *(Sars-Cov-2)* was already being considered an epidemic in the city of Wuhan in China $\!\!^1$, although it was difficult at first to identify the virus' origin, it's known that the Severe Acute Respiratory Syndrome *(SARS)* and the Middle East Respiratory Syndrome *(MERS)* usually need intermediate hosts $\!\!^2$, there is a possibility that bats and pangolins were the first carriers. After several public organizations called for caution relating to Covid-19, on March 12 of 2020 the pandemic state was recognized by the World Health Organization *(WHO)* $\!\!^3$.

 With the whole world being affected by the pandemic, it's a matter of great importance that studies about the spatio-temporal aspects of the Covid-19 pandemic should be conducted, the main goal of this paper is to present methods used to identify spatial and spatio-temporal dependence in the number of infections and deaths caused by Sars-Cov-2 in different countries accordingly to time in 2020, so these techniques may be used in future or hypothetical situations where is necessary to investigate the behavior of a pathogen agent in space and time.

 The spatial method was calculated by using the Global Moran's I and the Local Moran's I $\!\!^4$, which is also known as a LISA statistic (Local Indicator of Spatial Association), both being used as indicators of spatial association. Spatio-temporal association was evaluated by computing a spatio-temporal Poisson scan statistic, proposed by Kulldorff $\!\!^5$ to identify emerging or "alive" *clusters* in space and time.(later i'll add a few more things regarding two other studies i've found)

## Methods

 The statistical software R (Version 4.0.3) was used to calculate the spatial and the spatio-temporal association, the Covid-19 database used in the study is also available in R, in a package called "COVID19" built by the *COVID-19 Data Hub* $\!\!^1\!^0$, the database begins in January 2020 and is constantly updated, it also gathers data from various sources of information for all countries in which it was possible to obtain data. The database provides data about political measures adopted by each country during the pandemic, even though that information is not available for all countries included, which is one of the limitations of the database, that lack of data is also be seen in variables such as the number of recovered cases, where China for instance had no recovered cases during the analysis' period of time, since the reason for this was not explicit in the website of *COVID-19 Data Hub*, this paper supposes it was not possible to access data for countries which the information was lacking in the unified database.

 The cumulative number of deaths, infections and recovered cases is given by the database, but since these numbers are only available in its daily cumulative form, it was necessary to calculate monthly quantities for each one of the variables in the period of time considered in this paper between January and December 2020. The population of each country is also available in the database, so in order to better understand the spatial and the spatio-temporal effects of Covid-19 in each location, it was necessary to calculate the ratio between the number of infections, the number of recovered cases and the quantity of deaths accordingly to the population.

#### Moran's I

 To evaluate spatial association during the pandemic, a knn method was used to create spatial polygons and to build spatial weight matrices, which are necessary to calculate the Global Moran's I and the *LISA*. The Global Moran's I is widely used and gathers all spatial dependence in one value $\!\!^6$, while the use of the Local Moran's I is important to identify patterns of spatial association, hence the *LISA* should be used with the Global Moran's I, as it's stated by Maia, Gomes and Almeida (2019, p.930) "while the global Moran's I may suggest, in general, that there is little spatial autocorrelation in the data, LISA values can identify smaller geographic areas where positive or negative clustering occurs." $\!\!^7$.

 The Global Moran's I is calculated as shown in **(Equation 1)**, with $w$ as the spatial weight matrix, $x$ as the values of the studied variable, $n$ as the number of identified polygons and $\overline x$ as the mean or average:

**Equation 1**

$$Im = \frac{n\sum_{i=1}^{n}\sum_{j=1}^{n}(x_{i}-\overline{x})(x_{j}-\overline{x})w_{ij}}{W\sum_{i=1}^{n}(x_{i}-\overline{x})^2}$$
 Where W is defined by the following equation:
 $$W=\sum_{i=1}^{n}\sum_{j=1}^{n}w_{ij}$$

 The values range from -1 to 1, with the 0 value indicating an absence of spatial autocorrelation, positive values are indicators of positive spatial association, while negative values mean there is an inverse association, since this work seeks to find positive spatial association, the hypothesis tests were built within a 95% confidence level, so that the alternative hypothesis (H1) corresponds to an index higher than zero and the null hypothesis (Ho) to a lack of spatial association, both of which can be written mathematically as Ho=0 and H1>0. The monthly quantities were important to avoid problems of non-stationarity when computing the Global Moran's I, according to Cardoso (2007, p.3, translated from portuguese) "The index loses its validity when calculated for non-stationary data.", in order to validate the index 10,000 simulations were computed in a test of random permutations (*Monte-Carlo*) $\!\!^8$.

 Whereas the Local Moran's I (*LISA*) is calculated by the following equation (**Equation 2**):

**Equation 2**
 $$I(i) = \frac{x_{i}-\overline x}{s^{2}_{i}}\sum_{j=1}^{n}w_{ij}(x_{j}-\overline{x}), i=1,...,n, j\neq i$$
 Where $s^{2}_{i}$ is the variance of the studied variable $x_{i}$:
  $$s^{2}(i) = \frac{\sum_{j=1}^{n}(x_{j}-\overline x)^{2}}{n-1}, i=1,...,n, j\neq i$$

  Maps for the Local Moran's I, *Boxmaps* and Global Moran's I tables, were built for variables (number of cases and deaths) which achieved a global index IM>0.50. Boxmaps are cartographic representations of the Moran scatter plot, in which according to Seffrin, Araujo e Bazzi (2018,p.28)

   Os *Boxmaps* são uma representação cartográfica dos quadrantes do gráfico de dispersão de Moran, onde temos que em Seffrin, Araujo e Bazzi (2018,p.28) os valores AA e BB significam respectivamente que essas áreas possuem valores altos rodeadas por áreas que apresentam valores também altos e valores baixos da variável de interesse, rodeadas por áreas que apresentam valores baixos, o valor BA representa áreas com valor baixo rodeada de outras áreas com valores altos, já o AB é o contrário, são regiões com valores altos próximas de locais com valores baixos.

#### A estatística de Scan Poisson

 Em busca de identificar *clusters* emergentes no espaço-tempo, o método retrospectivo proposto por Kulldorff, aqui utilizada supondo uma distribuição de probabilidade poisson $\!\!^1\!^0$ (citação colocar traduzida kulldorff) "To detect emerging geographical clusters due to some suddenly occurring risk factor quickly can be of great public health importance. It can be used to detect previously unknown diseases, previously unknown risk factors for known diseases or the previously unknown local existence of known risk factors"(continuar escrevendo...)

## Resultados

 A partir do mês de Abril, o número acumulado de mortes por habitante apresentou uma correlação espacial no índice de Moran global (IM>0,50; P-valor<0,05) que se manteve em um alto patamar de correlação até o mês de Julho. Referente ao índice de Moran calculado para o número mensal de mortes por habitante, no mês de Abril o índice de Moran global (IM=0,72; P-valor<0,05) apresentou um alto patamar de dependência **(Tabela - 1)**, apresentando uma leve queda nos meses subsequentes em relação a Abril.

 Os índices locais de Moran para o número acumulado de mortes por habitante e para a quantidade mensal, dão um indicativo de uma grande dependência espacial entre a parte ocidental da Europa, América do Norte e América do Sul, havendo poucas diferenças entre os mapas de correlação local ao longo dos meses.

 A média móvel semanal da quantidade de casos por dia apresentou crescimento de Janeiro à Outubro, no mundo, na Europa e na América do Norte e Central, decaindo apenas na América do Sul, no entanto, a média semanal do número de mortes por dia apresentou uma variabilidade maior, porém com aumento em todas as localidades até o mês de outubro, sendo o aumento de maior impacto na Europa após uma brusca queda no número diário de falecidos pela Covid-19.

 Até o mês de Julho referente ao número cumulativo de morte por quantidade de casos o Iémen foi o país com maior número de mortos por infectados, chegando a um falecimento de 29% dos contagiados pelo Covid-19, no geral as Américas, a África, a Europa, a Ásia e a Oceania agiram de maneiras muito similares no quesito referente ao fechamento internacional de fronteiras no começo da pandemia. Dentre os continentes em que foi identificada dependência espacial sobre o número mensal de mortes por habitantes, de Fevereiro a Julho, houve similaridades nas políticas de rastreamento de casos, obrigação para ficar em casa e restrição de multidões.

## Discussão

(here)

## Conclusão

 (Here)

## Referências

[1] - Muhammad Adnan Shereen, Suliman Khan, Abeer Kazmi, Nadia Bashir, Rabeea Siddique, COVID-19 infection: Origin, transmission, and characteristics of human coronaviruses, Journal of Advanced Research, Volume 24, 2020, Pages 91-98, ISSN 2090-1232, https://doi.org/10.1016/j.jare.2020.03.005.

[2] - Tao Zhang, Qunfu Wu, Zhigang Zhang, Probable Pangolin Origin of SARS-CoV-2 Associated with the COVID-19 Outbreak, Current Biology, Volume 30, Issue 7, 2020, Pages 1346-1351.e2, ISSN 0960-9822, https://doi.org/10.1016/j.cub.2020.03.022.

[3] - WHO announces COVID-19 outbreak a pandemic. World Health Organization Regional Office for Europe. [S.l.]. Publicado em: 12/03/2020. <https://www.euro.who.int/en/health-topics/health-emergencies/coronavirus-covid-19/news/news/2020/3/who-announces-covid-19-outbreak-a-pandemic>. Acesso em: 24/11/2020.

[4] -CARDOSO, Carlos Eduardo de Paiva. Dependência Espacial, Setores Censitários, Zonas OD, Distritos, Sub Prefeituras e etc. [S.l]. Publicado em: 06/09/2007. <http://www.sinaldetransito.com.br/artigos/espacial.pdf>. Acesso em: 24/11/2020.

[10] - Guidotti, E., Ardia, D., (2020), “COVID-19 Data Hub”, Journal of Open Source Software 5(51):2376, doi: 10.21105/joss.02376.

[6] - BRAGA, Altemir da Silva. et al. Estudo de Dependência Espacial utilizando Análise de Dados de Área Aplicada na Mesorregião Metropolitana de Belo Horizonte por meio do Indicador Econômico PIB. 19ª SINAPE.  [S.l.]. 2010,  Disponível em: <http://www2.ime.unicamp.br/sinape/sites/default/files/Resumo%20expandido%20SINAPE.pdf>. Acesso em: 24/11/2020.

[7] - MAIA, André Luis Santiago; GOMES, Gecynalda Soares da Silva; ALMEIDA, Isabelle Galdino de. SPATIAL STUDY OF INCIDENCE RATES OF OCCUPATIONAL ACCIDENTS IN BRAZIL FROM 2002 TO 2012. REVISTA BRASILEIRA DE BIOMETRIA, [S.l.], v. 36, n. 4, p. 927-941, dec. 2018. ISSN 1983-0823. Available at: <http://www.biometria.ufla.br/index.php/BBJ/article/view/322>. Date accessed: 24 nov. 2020. doi: https://doi.org/10.28951/rbb.v36i4.322.

[8] - SEFFRIN, R.; ARAUJO, E. C.; BAZZI, C. L.. Análise espacial de área aplicada a produtividade de soja na região oeste do Paraná utilizando o software R. R. bras. Geom., Curitiba, v. 6, n. 1, p. 23-43, jan/mar.2018. Disponível em: <https://periodicos.utfpr.edu.br/rbgeo/article/view/5912>. Acesso em: 24/11/2020.DOI: 10.3895/rbgeo.v6n1.5912

[9] - Max Roser, Hannah Ritchie, Esteban Ortiz-Ospina and Joe Hasell (2020) - “Coronavirus Pandemic (COVID-19)”. Published online at OurWorldInData.org. Retrieved from: ‘https://ourworldindata.org/coronavirus’ [Online Resource]. Acesso em: 24/11/2020.

[5] - Kulldorff, M. (2001). Prospective time periodic geographical disease surveillance using a scan statistic. Journal of the Royal Statistical Society: Series A (Statistics in Society), 164(1), 61–72. doi:10.1111/1467-985x.00186

( Os países que não tiveram casos confirmados, ou que não estivessem no banco da *COVID-19 Data Hub*, não tiveram seus territórios presentes nos mapas.)-> vou usar